---
title: "Overview of variables"
description: |
  The goal is to generate a table that allows the reader to search for variables by their acronym, get an explanation what a variable measures as well as mean and standard deviation. I want to produce a table with more information on each variable so that you can just easily look up a variable. Such a table would be for everyone interested in the data set.
author:
  - name: Sophie Pychlau
    url: https://github.com/spychlau
date: 03-16-2020
output:
  distill::distill_article:
    self_contained: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  comment = "##",
  layout = "l-body-outset",
  R.options = list(width = 60),
  fig.height = 10,
  fig.width = 10
)

library(tidyverse)
library(reactable)
library(gt)

variable_names <- read_csv("variables2.csv") 

variable_names <- variable_names %>% 
  mutate(Function = as.factor(Function))

data <- read_csv("data_syn.csv") 

```

Using the gt package seems a natural first step:

```{r}

variable_names %>% 
  gt()

```

This is a nice and easy overview of the variables I used, quickly generated based on an excel file that contained all this information. However, R offers me the opportunity to add more information to this table. I can easily compute some simple statistics and join them to the file. 

<aside>
Keep in mind that all variables (except vofrequ, a categorical variable with 7 levels) were measured on a Likert scale from 1 ("strongly disagree") to 5 ("strongly agree").
</aside>

```{r}

descriptives <- data %>% 
  psych::describe() %>% 
  mutate(Acronym = as.factor(vars))

#The new file with the descriptives does not contain the acronyms. Before merging, I therefore need to add that column:
levels(descriptives$Acronym) <-  c("prhvos", "prmvos", "morobs", "guilt", "pride", "emorep", "vofrequ", "prhvoss", "obse", 
                                   "morobss", "feeltr", "SER", "selfas", "VoicSat", "frame", "comval", "authov", "corese", 
                                   "ueprorg", "shame", "constr")

variables <- variable_names %>% 
  left_join(descriptives, by = "Acronym") %>% 
  mutate(Mean = round(mean, 2)) %>% 
  mutate(SD = round(sd, 2))

#I reorder the columns in the order I want them displayed
variables <- variables[, c(1, 3, 2, 18, 19, 4)]

variables %>% 
  gt()

```

Using gt is a very easy and attractive way of making a table. However, given that the purpose of this blog is to provide readers with a way to search for variables, an interactive table would be superior to gt. Therefore, reactable is good:

```{r, layout = "l-page"}

reactable(variables,
          resizable = TRUE,
          striped = TRUE,
          highlight = TRUE,
          searchable = TRUE)

```

These are the interactive features I want! Now I just want to make the table look nicer. In particulat, I want adjust the column width and order the first column alphabetically by default.

<aside>
Specifying the width of the columns was just guess work.
</aside>

```{r, layout = "l-page"}

variables %>%
  reactable(
    searchable = TRUE,
    striped = TRUE,
    highlight = TRUE,
    pagination = FALSE,
    defaultSorted = list(Acronym = "asc"),
   columns = list(Explanation = 
      colDef(minWidth = 200),
      Function = 
      colDef(maxWidth = 100),
      Mean = 
      colDef(maxWidth = 70),
      SD = 
      colDef(maxWidth = 70),
      Acronym = 
      colDef(maxWidth = 90)))

```


